A Dependently Typed Multi-Stage Calculus \ifnotanonymous\thanks{Supported by organization x.}\fi

### 2019-06-15 text/formal-defnition.tex
Wrote on vadd and composition.

%%%%%%%%%%%%%%%%%%% SKIP FROM HERE %%%%%%%%%%%%%%%%%%%%
$\text{mulmat}$ function in the following code fragment generate code for matrix multiplication.
$\text{mulmat}$ takes two integers which are the size of the multiplicand matrix and generate a code.
The last integer to decide the size of the multiplier matrix is given at runtime.
You can generate code by applying two integers to $\text{mulmat}$.
We applied 3 and 5 in the second line and got $\TW_\alpha \Pi z:\I.(\M\ z\ \%_\alpha 5) \to (\M\ \%_\alpha 5\ \%_\alpha 3) \to (\M\ z\ \%_\alpha 3)$.
But this type is difficult to combine with other code because there are two $\%_\alpha$ for CSP.
\QPercent\ states that we can erase this $\%_\alpha$ under a condition which is explained in the next paragraph.

The condition is that a CSPed value equals to the original value when it has the same type in the original stage.
For example, $\V \%_\alpha 5 \E 5 @ \alpha$ because $ \V 5 : \I @ \alpha $ from \TConst\ and  $ \V \%_\alpha 5 : \I @ \alpha$.
In other words, we can remove a $\%_\alpha$ symbol of a value when it doesn't change the type.

    % Example of QCsp (mulmat)
    {
        \begin{align*}
            \text{mulmat}       & : \Pi x:\I.\Pi y:\I.(\TW_\alpha \Pi z:\I.(\M\ z\ \%_\alpha y) \to (\M\ \%_\alpha y\ \%_\alpha x) \to (\M\ z\ \%_\alpha x)) \\
            \text{mulmat}\ 3\ 5 & : \TW_\alpha \Pi z:\I.(\M\ z\ \%_\alpha 5) \to (\M\ \%_\alpha 5\ \%_\alpha 3) \to (\M\ z\ \%_\alpha 3)                     \\
                                & (\E \TW_\alpha \Pi z:\I.(\M\ z\ 5) \to (\M\ 5\ 3) \to (\M\ z\ 3) )                                                         \\
        \end{align*}
    }

This kind of type equivalence is very useful when we combine $\text{mulmat}\ 3\ 5$ with another code.
We cannot remove $\%_\alpha$ in the type without \QPercent and it makes very difficult to combine
because types of codes don't contain $\%_\alpha$ generally.

%%%%%%%%%%%%%%%%%%% TO HERE %%%%%%%%%%%%%%%%%%%%

>>>>

\renewcommand{\Vpn}{\text{Vector}\ (\%_\alpha n)}
As we discussed in Section 2, $\text{vadd}$ function, 
which takes the length of vectors and returns a quoted add function for vector of the given size,
has the type of $\Pi n:\I.\F\beta.\TW_\beta(\Vpn\to\Vpn\to\Vpn)$.
When we use this function, we instansiate $\text{vadd}$ with some integer such as 5
and get code of a optimzed $\text{vadd}$ function such as $\text{vadd5}$.
\renewcommand{\Vpn}{\text{Vector}\ (\%_\beta 5)}
Although $\text{vadd5}$ is of type $\F\beta.\TW_\beta(\Vpn\to\Vpn\to\Vpn)$,
a code value of this type is difficult to combine with another code fragment.
For example, if we want to composite $\text{vadd5}$ with reverse fucntion, 
which is of type $\text{rev5}:\F\alpha.\TW_\alpha(\text{Vector}\ 5\to\text{Vector}\ 5)$,
we write a program like following.
\begin{align*}
\Lambda\gamma.\TB_\gamma.\lambda v_1:(\text{Vector}\ 5).\lambda v_2:(\text{Vector}\ 5).\TBL_\gamma 
(\text{rev5}\ \gamma)\ (\TBL_\gamma (\text{vadd5}\ \gamma)\ \TBL_\gamma v_1\ \TBL_\gamma v_2)
\end{align*}
However, this program cannot be typed without \QPercent{}
because the type of $v_1$: $\text{Vector}\ 5$ isn't equivalent with the type of arguments of $\text{vadd5}$: $\text{Vector}\ \%_\gamma 5$.

In order to resolve this problem, we introduce \QPercent{} into \LMD.
\begin{center}
\infrule[\QPercent]{\G\V M:\tau @{A\alpha} \andalso \G\V M:\tau @A}{\G\V\%_\alpha M \E M : \tau @{A\alpha}}
\end{center}
\QPercent\ states that we can erase this $\%_\alpha$ under a condition.
The condition is that a CSPed value equals to the original value when it has the same type in the original stage.
For example, $\V \%_\alpha 5 \E 5 @ \alpha$ because $ \V 5 : \I @ \alpha $ from \TConst\ and  $ \V \%_\alpha 5 : \I @ \alpha$.
In other words, we can remove a $\%_\alpha$ symbol of a value when it doesn't change the type.
Thanks to \QPercent, we can identify $\text{Vector}\ 5$ and $\text{Vector}\ \%_\gamma 5$
and compose code values of functions as we expected.
